---
title: "20240904_lab0"
author: "Yerim Oh"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Beer example

In this lab, "we examine data from sixteen student volunteers at Ohio State University who each drank a randomly assigned number of cans of beer. These students were evenly divided between men and women, and they differed in weight and drinking habits. Thirty minutes later, a police officer measured their blood alcohol content (BAC) in grams of alcohol per deciliter of blood."[1,2]

References:

(1) Diez, David M., Christopher D. Barr, and Mine Cetinkaya-Rundel. OpenIntro statistics. Boston, MA, USA. OpenIntro, 2012.
(2) J. Malkevitch and L.M. Lesser. For All Practical Purposes: Mathematical Literacy in Today's World. WH Freeman & Co, 2008.

```{r, echo= FALSE}
#The data:
beer <- structure(list(beers = c(5L,2L,9L,8L,3L,7L,3L,5L,3L,5L,4L,6L,5L,7L,1L,4L),
                       bac = c(0.10000000000000001,0.029999999999999999, 0.19, 0.12,
                               0.040000000000000001, 0.095000000000000001, 
                               0.070000000000000007, 0.059999999999999998, 0.02,
                               0.050000000000000003, 0.070000000000000007,
                               0.10000000000000001, 0.085000000000000006,
                               0.089999999999999997, 0.01, 0.050000000000000003)),
                  row.names = c(NA, -16L), class = c("tbl_df", "tbl", "data.frame"))
```

### a) What are the response and explanatory variables?

response variable: blood alcohol content (BAC) in grams of alcohol per deciliter of blood

explanatory variable: number of cans of beer

### b) Plot the data and describe the relationship between the variables.

```{r, error = TRUE}
ggplot(data = beer, mapping = aes(x = beers, y = bac)) + 
  geom_point() +
  theme_bw()
```

### c) Which function $f$ could describe the relationship between $X$ and $Y$?

A linear relationship seems appropriate.

Estimated equation of the line: $f$(beers) $=\beta_0 + \beta_1 \times$ beers

### d) How could we estimate this function (name the method or code it in the following R chunk)?

Method of Least Squares

$\hat{f}(\text{beers}) = -0.0127+0.01796 \times \text{beers}$

```{r}
# fit the model
beer_model <- lm(bac ~ beers, data = beer)
summary(beer_model)

# plot the linear function: 2 ways
plot(beer$beers, beer$bac,
     xlab = "num of cans", ylab = "BAC",
     pch = 19, frame = FALSE)
abline(beer_model, col = "blue")

beer <- beer %>%
  mutate(fitted = predict(beer_model))
ggplot() +
  geom_point(data = beer, mapping = aes(x = beers, y = bac)) +
  geom_line(data = beer, mapping = aes(x = beers, y = fitted), color = "blue") +
  theme_bw()
```
